---
title: "Epidemiological Surveillance Report"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  word_document:
    toc: false
params:
  data: NULL
  indicators: NULL
  period: "weekly"
  geo_level: "district"
  title: "Epidemiological Surveillance Report"
  location_name: "All locations"
  logo_path: NULL
  report_date: !r Sys.Date()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 7,
  fig.height = 4
)
library(dplyr)
library(ggplot2)
library(flextable)
library(lubridate)

data       <- params$data
indicators <- params$indicators
loc_name   <- params$location_name
rep_date   <- params$report_date
title      <- params$title

date_min <- format(min(data$date), "%d/%m/%Y")
date_max <- format(max(data$date), "%d/%m/%Y")
week_num <- isoweek(max(data$date))
year_num <- year(max(data$date))

total_cases  <- sum(indicators$total_cases, na.rm = TRUE)
total_deaths <- if ("total_deaths" %in% names(indicators)) sum(indicators$total_deaths, na.rm = TRUE) else NA
```

# `r title`

**Location:** `r loc_name` | **Week:** `r week_num` – `r year_num` | **Period:** `r date_min` to `r date_max`

---

## 1. Executive Summary

```{r summary-table}
n_diseases  <- n_distinct(data$disease)
n_locations <- n_distinct(data$location)

summary_df <- data.frame(
  Indicator = c(
    "Reporting period",
    "Diseases under surveillance",
    "Geographic units",
    "Total cases reported",
    "Total deaths reported",
    "Overall CFR (%)"
  ),
  Value = c(
    paste0(date_min, " to ", date_max),
    as.character(n_diseases),
    as.character(n_locations),
    as.character(total_cases),
    if (!is.na(total_deaths)) as.character(total_deaths) else "N/A",
    if (!is.na(total_deaths) && total_cases > 0)
      paste0(round(total_deaths / total_cases * 100, 1), "%")
    else "N/A"
  )
)

flextable(summary_df) |>
  set_header_labels(Indicator = "Indicator", Value = "Value") |>
  bold(part = "header") |>
  bg(part = "header", bg = "#2166ac") |>
  color(part = "header", color = "white") |>
  autofit() |>
  border_outer() |>
  border_inner_h()
```

---

## 2. Case Summary by Disease

```{r disease-table}
disease_summary <- indicators |>
  group_by(disease) |>
  summarise(
    Total_Cases  = sum(total_cases, na.rm = TRUE),
    Total_Deaths = if ("total_deaths" %in% names(indicators))
                     sum(total_deaths, na.rm = TRUE) else NA_real_,
    CFR_pct      = if (!is.na(Total_Deaths) && Total_Cases > 0)
                     round(Total_Deaths / Total_Cases * 100, 1) else NA_real_,
    .groups = "drop"
  ) |>
  arrange(desc(Total_Cases))

names(disease_summary) <- c("Disease", "Total Cases", "Total Deaths", "CFR (%)")

flextable(disease_summary) |>
  bold(part = "header") |>
  bg(part = "header", bg = "#2166ac") |>
  color(part = "header", color = "white") |>
  bg(i = seq(2, nrow(disease_summary), by = 2), bg = "#f0f4f8") |>
  autofit() |>
  border_outer() |>
  border_inner_h()
```

---

## 3. Epidemic Curves

```{r epi-curves, fig.cap="Weekly epidemic curves by disease"}
plot_data <- data |>
  mutate(week_date = floor_date(date, unit = "week")) |>
  group_by(week_date, disease) |>
  summarise(cases = sum(cases, na.rm = TRUE), .groups = "drop")

ggplot(plot_data, aes(x = week_date, y = cases)) +
  geom_col(fill = "#2166ac", color = "white", width = 6) +
  facet_wrap(~ disease, scales = "free_y") +
  scale_x_date(date_labels = "W%V\n%Y", date_breaks = "1 week") +
  labs(
    title    = paste0("Epidemic Curves — Week ", week_num, ", ", year_num),
    subtitle = loc_name,
    x        = "Epidemiological Week",
    y        = "Number of Cases"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40"),
    strip.text    = element_text(face = "bold"),
    axis.text.x   = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )
```

---

## 4. Weekly Indicators

```{r indicators-table}
ind_display <- indicators |>
  select(
    period_label, disease, location, total_cases,
    dplyr::any_of(c("incidence_per_100k", "total_deaths", "cfr_pct"))
  ) |>
  arrange(period_label, disease, location)

col_labels <- c(
  period_label       = "Week",
  disease            = "Disease",
  location           = "Location",
  total_cases        = "Cases",
  incidence_per_100k = "Incidence(/100k)",
  total_deaths       = "Deaths",
  cfr_pct            = "CFR (%)"
)

flextable(ind_display) |>
  set_header_labels(values = col_labels[names(col_labels) %in% names(ind_display)]) |>
  bold(part = "header") |>
  bg(part = "header", bg = "#2166ac") |>
  color(part = "header", color = "white") |>
  bg(i = seq(2, nrow(ind_display), by = 2), bg = "#f0f4f8") |>
  autofit() |>
  border_outer() |>
  border_inner_h()
```

---

## 5. Alerts

```{r alerts}
if ("cfr_pct" %in% names(indicators)) {
  alerts <- indicators |>
    filter(!is.na(cfr_pct) & cfr_pct > 1) |>
    select(disease, location, period_label, cfr_pct) |>
    arrange(desc(cfr_pct))

  if (nrow(alerts) > 0) {
    cat("**ALERT — High Case Fatality Rate detected:**\n\n")
    for (i in seq_len(nrow(alerts))) {
      cat(paste0("- **", alerts$disease[i], "** in ", alerts$location[i],
                 " (", alerts$period_label[i], "): CFR = ",
                 alerts$cfr_pct[i], "%\n"))
    }
  } else {
    cat("No CFR alerts this week.\n")
  }
}
```

---

## 6. Recommendations

- Continue routine surveillance for all notifiable diseases.
- Investigate any clusters exceeding alert thresholds.
- Ensure timely reporting from all health facilities.
- Share this bulletin with district health teams and partners.

---

*Report generated automatically by **epiforge** v0.1.0 on `r format(Sys.Date(), "%d %B %Y")`.*
